---
title: "curation"
author: "Sofia Salazar"
date: "5/16/2022"
output: html_document
---

## Curation of datasets

For each experiment a new column `DISEASE` is added in order to get uniform information about the disease status of each sample, this is done manually for each study to account for each author's way of classifiying their samples.

```{bash}
qlogin
cd /mnt/Citosina/amedina/ssalazar/meta/out
module load r/4.0.2
R
```

## Libraries

```{r}
library("recount3")
```


```{r}
SRA_projects <- scan(file = "/mnt/Citosina/amedina/ssalazar/meta/data/SRA-IDs.txt", character(), sep=",")
curation <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/downloads.xz")
names(curation) # we can see the studies downloaded
```

## Information curation

We will add a new column and use controlled vocabulary, so we will change the 
description that each authors used to describe SLE or control samples.

## `r names(curation)[1] `

```{r}
i <- 1
rse_object <- curation[i]
print(SRA_ID <- SRA_projects[i])
rse_gene <- rse_object$SRP168421

# identify the way authors described SLE or CONTROLS
rse_gene$sra_attribute.disease
rse_gene$sra_attribute.disease <- as.factor(rse_gene$sra_attribute.disease)

# creating a DISEASE column
rse_gene$DISEASE <- factor(ifelse(rse_gene$sra_attribute.disease == "systemic lupus erythematosus", "SLE", "CONTROL"))
table(rse_gene$DISEASE)

# overwritting curated rse object
curation[i] <- rse_gene
```

## `r names(norm)[2] `

```{r}
i <- 2
rse_object <- curation[i]
print(SRA_ID <- SRA_projects[i])
rse_gene <- rse_object$SRP062966

# identify the way authors described SLE or CONTROLS
rse_gene$sra_attribute.disease_status
rse_gene$sra_attribute.disease <- as.factor(rse_gene$sra_attribute.disease_status)

# creating a DISEASE column
rse_gene$DISEASE <- factor(ifelse(rse_gene$sra_attribute.disease_status == "systemic lupus erythematosus (SLE)", "SLE", "CONTROL"))
table(rse_gene$DISEASE)

# overwritting curated rse object
curation[i] <- rse_gene
```

## Output

```{r}
saveRDS(curation, file = "/mnt/Citosina/amedina/ssalazar/meta/out/curated_rse.xz", compress="xz")
```

## Session Info

```{r}
sessionInfo()
# R version 4.0.2 (2020-06-22)
# Platform: x86_64-pc-linux-gnu (64-bit)
# Running under: CentOS Linux 7 (Core)

# Matrix products: default
# BLAS:   /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRblas.so
# LAPACK: /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRlapack.so

# locale:
 # [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
 # [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
 # [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 # [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
 # [9] LC_ADDRESS=C               LC_TELEPHONE=C            
# [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

# attached base packages:
# [1] parallel  stats4    stats     graphics  grDevices utils     datasets 
# [8] methods   base     

# other attached packages:
 # [1] recount3_1.0.7              SummarizedExperiment_1.20.0
 # [3] Biobase_2.50.0              GenomicRanges_1.42.0       
 # [5] GenomeInfoDb_1.26.7         IRanges_2.24.1             
 # [7] S4Vectors_0.28.1            BiocGenerics_0.36.1        
 # [9] MatrixGenerics_1.2.1        matrixStats_0.60.0         

# loaded via a namespace (and not attached):
 # [1] tidyselect_1.1.0         purrr_0.3.4              lattice_0.20-41         
 # [4] vctrs_0.3.8              generics_0.1.0           BiocFileCache_1.14.0    
 # [7] rtracklayer_1.50.0       utf8_1.2.2               blob_1.2.1              
# [10] XML_3.99-0.6             rlang_0.4.11             R.oo_1.24.0             
# [13] pillar_1.6.2             withr_2.4.2              glue_1.4.2              
# [16] DBI_1.1.1                R.utils_2.10.1           BiocParallel_1.24.1     
# [19] rappdirs_0.3.3           bit64_4.0.5              dbplyr_2.0.0            
# [22] sessioninfo_1.1.1        GenomeInfoDbData_1.2.4   lifecycle_1.0.0         
# [25] zlibbioc_1.36.0          Biostrings_2.58.0        R.methodsS3_1.8.1       
# [28] memoise_2.0.0            fastmap_1.1.0            curl_4.3.2              
# [31] fansi_0.5.0              Rcpp_1.0.7               cachem_1.0.5            
# [34] DelayedArray_0.16.3      XVector_0.30.0           bit_4.0.4               
# [37] Rsamtools_2.6.0          dplyr_1.0.3              grid_4.0.2              
# [40] cli_3.0.1                tools_4.0.2              bitops_1.0-7            
# [43] magrittr_2.0.1           RCurl_1.98-1.3           RSQLite_2.2.7           
# [46] tibble_3.1.3             crayon_1.4.1             pkgconfig_2.0.3         
# [49] ellipsis_0.3.2           Matrix_1.3-4             data.table_1.14.0       
# [52] rstudioapi_0.13          assertthat_0.2.1         httr_1.4.2              
# [55] R6_2.5.0                 GenomicAlignments_1.26.0 compiler_4.0.2    
```

