---
title: "curation"
author: "Sofia Salazar"
date: "5/16/2022"
output: html_document
---

## Curation of datasets

For each experiment a new column `DISEASE` is added in order to get uniform information about the disease status of each sample, this is done manually for each study to account for each author's way of classifiying their samples.

```{bash}
cd /mnt/Citosina/amedina/ssalazar/meta/data
# SRP111941
awk -F '\t' -v OFS='\t' '{print $1,$25}' SraRunTable-SRP111941.tsv > info-SRP111941.tsv
# SRP296987
awk -F '\t' -v OFS='\t' '{print $1,$25}' SraRunTable-SRP296987.tsv > info-SRP296987.tsv
# SRP311059
awk -F '\t' -v OFS='\t' '{print $1,$14}' SraRunTable-SRP311059.tsv > info-SRP311059.tsv
# SRP322015
awk -F '\t' -v OFS='\t' '{print $1,$24}' SraRunTable-SRP322015.tsv > info-SRP322015.tsv
# SRP073191
awk -F '\t' -v OFS='\t' '{print $1,$13}' SraRunTable-SRP073191.tsv > info-SRP073191.tsv
# SRP136102
awk -F '\t' -v OFS='\t' '{print $1,$13}' SraRunTable-SRP136102.tsv > info-SRP136102.tsv

```

```{bash}
qlogin
cd /mnt/Citosina/amedina/ssalazar/meta/out
module load r/4.0.2
R
```

## Libraries

```{r}
library("recount3")
library(tidyverse)
```


```{r}
SRA_projects <- scan(file = "/mnt/Citosina/amedina/ssalazar/meta/data/SRA-IDs.txt", character(), sep=",")
curation <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/downloads.xz")
names(curation) # we can see the studies downloaded
```

## Information curation

Function to obtain metadata information of each study.

```{r}
obtainInfo <- function(df, rse_gene, name){
  df<- df %>%add_row(
    "SRA" = name,
    "CONTROL" = as.numeric(table(rse_gene$DISEASE)[1]),
    "SLE" = as.numeric(table(rse_gene$DISEASE)[2]),
    "NUM_GENES" = as.numeric(table((assays(rse_gene)$counts != 0))["TRUE"])
  )
}
```

Function to create a tibble with metadata information of each study.

```{r}
######## SUMMARY TABLE
## Summary information about the experiment
df <- tibble("SRA"= NA,
             "CONTROL" = NA,
             "SLE" = NA,
             "NUM_GENES" = NA
             #"H_AGE_MEAN" = NA,
             #"H_AGE_MEDIAN" = NA,
             #"H_AGE_RANGE" = NA,
             #"H_SEX_F" = NA,
             #"H_SEX_M" = NA,
             #"SLE_AGE_MEAN" = NA,
             #"SLE_AGE_MEDIAN" = NA,
             #"SLE_AGE_RANGE" = NA,
             #"SLE_SEX_F" = NA,
             #"SLE_SEX_M" = NA,
                 )
```

We will add a new column and use controlled vocabulary, so we will change the 
description that each authors used to describe SLE or control samples.

```{r}
# SRP168421
i <- 7
rse_gene <- curation[[i]]

# identify the way authors described SLE or CONTROLS
rse_gene$sra_attribute.disease
rse_gene$sra_attribute.disease <- as.factor(rse_gene$sra_attribute.disease)

# creating a DISEASE column
rse_gene$DISEASE <- factor(ifelse(rse_gene$sra_attribute.disease == "systemic lupus erythematosus", "SLE", "CONTROL"))
table(rse_gene$DISEASE)

df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))


# overwritting curated rse object
curation[i] <- rse_gene
```


```{r}
# SRP062966
i <- 8
rse_gene <- curation[[i]]

# identify the way authors described SLE or CONTROLS
rse_gene$sra_attribute.disease_status
rse_gene$sra_attribute.disease <- as.factor(rse_gene$sra_attribute.disease_status)

# creating a DISEASE column
rse_gene$DISEASE <- factor(ifelse(rse_gene$sra_attribute.disease_status == "systemic lupus erythematosus (SLE)", "SLE", "CONTROL"))
table(rse_gene$DISEASE)

df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))


# overwritting curated rse object
curation[i] <- rse_gene
```


```{r}
# SRP073191
i <- 5
rse_gene <- curation[[i]]

# identify the way authors described SLE or CONTROLS
rse_gene$sra_attribute.disease_state
rse_gene$sra_attribute.disease_state <- as.factor(rse_gene$sra_attribute.disease_state)

# creating a DISEASE column
rse_gene$DISEASE <- factor(ifelse(rse_gene$sra_attribute.disease_state == "Systemic lupus erythematosus", "SLE", "CONTROL"))
table(rse_gene$DISEASE)

df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))


# overwritting curated rse object
curation[i] <- rse_gene
```

```{r}
# SRP136102
i <- 6
rse_gene <- curation[[i]]

# identify the way authors described SLE or CONTROLS
rse_gene$sra_attribute.phenotype
rse_gene$sra_attribute.phenotype <- as.factor(rse_gene$sra_attribute.phenotype)

# creating a DISEASE column
rse_gene$DISEASE <- factor(ifelse(rse_gene$sra_attribute.phenotype == "SLE", "SLE", "CONTROL"))
table(rse_gene$DISEASE)

df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))


# overwritting curated rse object
curation[i] <- rse_gene
```

```{r}
# SRP111941
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP111941.tsv", header = T, sep= '\t')
i <- 1
sra_info$subject_status <- as.factor(sra_info$subject_status)
rse_gene <- curation[[i]]
rse_gene$DISEASE <- factor(ifelse(sra_info$subject_status == "SLE patient", "SLE", "CONTROL"))
table(rse_gene$DISEASE)
df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))
# overwritting curated rse object
curation[i] <- rse_gene
```

```{r}
# SRP296987
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP296987.tsv", header = T, sep= '\t')
i <- 2
sra_info$source_name <- as.factor(sra_info$source_name)
rse_gene <- curation[[i]]
rse_gene$DISEASE <- factor(ifelse(sra_info$source_name == "systemic lupus erythematosus", "SLE", "CONTROL"))
table(rse_gene$DISEASE)
df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))
# overwritting curated rse object
curation[i] <- rse_gene
```

```{r}
# SRP311059
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP311059.tsv", header = T, sep= '\t')
i <- 3
sra_info$Diagnosis <- as.factor(sra_info$Diagnosis)
rse_gene <- curation[[i]]
rse_gene$DISEASE <- factor(ifelse(sra_info$Diagnosis == "SLE patient", "SLE", "CONTROL"))
table(rse_gene$DISEASE)
df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))
# overwritting curated rse object
curation[i] <- rse_gene
```

```{r}
# SRP322015
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP322015.tsv", header = T, sep= '\t')
i <- 4
sra_info$source_name <- as.factor(sra_info$source_name)
rse_gene <- curation[[i]]
rse_gene$DISEASE <- factor(ifelse(sra_info$source_name == "Tibetan SLE patients", "SLE", "CONTROL"))
table(rse_gene$DISEASE)
df<- obtainInfo(df, rse_gene = rse_gene, name=names(curation[i]))
# overwritting curated rse object
curation[i] <- rse_gene
```


## Output

```{r}
saveRDS(curation, file = "/mnt/Citosina/amedina/ssalazar/meta/out/curated_rse.xz", compress="xz")
df <- df[-1,]
write_csv(df, file = "/mnt/Citosina/amedina/ssalazar/meta/out/summary.csv")
```

## Session Info

```{r}
sessionInfo()
# R version 4.0.2 (2020-06-22)
# Platform: x86_64-pc-linux-gnu (64-bit)
# Running under: CentOS Linux 7 (Core)

# Matrix products: default
# BLAS:   /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRblas.so
# LAPACK: /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRlapack.so

# locale:
 # [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
 # [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
 # [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 # [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
 # [9] LC_ADDRESS=C               LC_TELEPHONE=C            
# [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

# attached base packages:
# [1] parallel  stats4    stats     graphics  grDevices utils     datasets 
# [8] methods   base     

# other attached packages:
 # [1] forcats_0.5.1               stringr_1.4.0              
 # [3] dplyr_1.0.3                 purrr_0.3.4                
 # [5] readr_1.4.0                 tidyr_1.1.2                
 # [7] tibble_3.1.3                ggplot2_3.3.5              
 # [9] tidyverse_1.3.0             recount3_1.0.7             
# [11] SummarizedExperiment_1.20.0 Biobase_2.50.0             
# [13] GenomicRanges_1.42.0        GenomeInfoDb_1.26.7        
# [15] IRanges_2.24.1              S4Vectors_0.28.1           
# [17] BiocGenerics_0.36.1         MatrixGenerics_1.2.1       
# [19] matrixStats_0.60.0         

# loaded via a namespace (and not attached):
 # [1] httr_1.4.2               bit64_4.0.5              jsonlite_1.7.2          
 # [4] R.utils_2.10.1           modelr_0.1.8             assertthat_0.2.1        
 # [7] BiocFileCache_1.14.0     blob_1.2.1               cellranger_1.1.0        
# [10] GenomeInfoDbData_1.2.4   Rsamtools_2.6.0          sessioninfo_1.1.1       
# [13] pillar_1.6.2             RSQLite_2.2.7            backports_1.2.1         
# [16] lattice_0.20-41          glue_1.4.2               XVector_0.30.0          
# [19] rvest_0.3.6              colorspace_2.0-2         Matrix_1.3-4            
# [22] R.oo_1.24.0              XML_3.99-0.6             pkgconfig_2.0.3         
# [25] broom_0.7.9              haven_2.3.1              zlibbioc_1.36.0         
# [28] scales_1.1.1             BiocParallel_1.24.1      generics_0.1.0          
# [31] ellipsis_0.3.2           cachem_1.0.5             withr_2.4.2             
# [34] cli_3.0.1                readxl_1.3.1             magrittr_2.0.1          
# [37] crayon_1.4.1             memoise_2.0.0            R.methodsS3_1.8.1       
# [40] fs_1.5.0                 fansi_0.5.0              xml2_1.3.2              
# [43] tools_4.0.2              data.table_1.14.0        hms_1.1.0               
# [46] lifecycle_1.0.0          reprex_1.0.0             munsell_0.5.0           
# [49] DelayedArray_0.16.3      Biostrings_2.58.0        compiler_4.0.2          
# [52] rlang_0.4.11             grid_4.0.2               RCurl_1.98-1.3          
# [55] rstudioapi_0.13          rappdirs_0.3.3           bitops_1.0-7            
# [58] gtable_0.3.0             DBI_1.1.1                curl_4.3.2              
# [61] R6_2.5.0                 GenomicAlignments_1.26.0 lubridate_1.7.9.2       
# [64] rtracklayer_1.50.0       fastmap_1.1.0            bit_4.0.4               
# [67] utf8_1.2.2               stringi_1.6.2            Rcpp_1.0.7              
# [70] vctrs_0.3.8              dbplyr_2.0.0             tidyselect_1.1.0  
```

