---
title: "DE"
author: "Sofia Salazar"
date: "5/16/2022"
output: html_document
---

## Setup

This script is used to perform a DE analysis in the previously normalized `DGEList` objects obtained from the previous `normalization-3.Rmd` script.

```{bash}
qlogin
cd /mnt/Citosina/amedina/ssalazar/meta/out
module load r/4.0.2
R
```

## Libraries

```{r}
library("recount3")
library(edgeR)
library(tidyverse)
library(limma)
library(DESeq2)
```

## Differential expression

Getting the normalized `DGEList` objects obtained from previous `normalization.Rmd` and the `rse` objects obtained from the previous `curation.Rmd` script.


```{r}
SRA_projects <- scan(file = "/mnt/Citosina/amedina/ssalazar/meta/data/SRA-IDs.txt", character(), sep=",")
# DE_analysis <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/normalized_dge.xz")
dge_1 <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/SRP296987_dge.xz")
dge_2 <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/SRP062966_dge.xz")
dge_3 <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/SRP073191_dge.xz")
dge_4 <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/SRP136102_dge.xz")
rse_curated <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/corrected_rse.xz")
# copy to overwrite
DE_analysis <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/corrected_rse.xz")
```

The following function performs the DE analysis for each `DGEList` given as input.

```{r}
do_limma <- function(dge, rse_gene){
  design <- model.matrix(as.formula(~ DISEASE), data = colData(rse_gene))
  # "When the library sizes are quite variable between samples, then the voom approach is theoretically more powerful than limma-trend."
  vGene <- voom(dge, design, plot = F)
  # eBayes in lmFit model
  eb_results <- eBayes(lmFit(vGene))
  print(colnames(coef(lmFit(vGene))))
  # volcanoplot(eb_results,coef = 2, highlight=20, pch=20)
  de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_gene), sort.by = "none", confint = T)
  # formatting it as a tibble
  res <- as_tibble(de_results,rownames="rownames")
}
```

```{r}
 do_DESeq <-function(rse){ 
  dds <- DESeqDataSet(rse, ~DISEASE)
  keep <- rowSums(counts(dds)) >= 10
  dds <- dds[keep,]
  dds <- DESeq(dds, test = "Wald")
  res <- results(dds)
  res <- as_tibble(res, rownames="rownames")
 }
```

Applying the functions

```{r}
#for(i in 1:length(SRA_projects)){
  #rse_object <- rse_curated[i]
  #rse <- rse_object[[1]]
  #dge_object <- DE_analysis[i]
  # apply the function to the rse object inside each list name
  #DE_result <- lapply(dge_object, do_DE, rse_gene = rse)
  #DE_bayes_result <- lapply(rse, do_DE_bayes)
  #DE_bayes_result <- lapply(dge_object, do_DE_bayes, rse_gene = rse)
  # renaming colnames with the SRA id
  #colnames(DE_result[[1]]) <- str_c(colnames(DE_result[[1]]),"_", names(rse_curated[i]))
  
  #colnames(DE_bayes_result[[1]]) <- str_c(colnames(DE_bayes_result[[1]]),"_", names(rse_curated[i]))
   # saving as csv
  #write_csv(as.data.frame(DE_result[[1]]), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/DE/Table_Genes_", names(rse_curated[i]), ".csv"))
  #write_csv(as.data.frame(DE_bayes_result[[1]]), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/DE_bayes/Table_Genes_", names(rse_curated[i]), ".csv"))
  # overwriting DE_analysis[i] to use later
  #DE_analysis[i] <- DE_result
#}
```

Limma for SRP296987, SRP062966, SRP073191, SRP136102

```{r}
DE <- do_limma(dge_1, rse_gene = rse_curated$SRP296987)
DE$lfcSE <- (DE$CI.R-DE$CI.L)/(2*1.96)
DE_analysis$SRP296987 <- DE
colnames(DE_analysis$SRP296987) <- str_c(colnames(DE_analysis$SRP296987),"_SRP296987")
write_csv(as.data.frame(DE_analysis$SRP296987), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP296987.csv")

DE <- do_limma(dge_2, rse_gene = rse_curated$SRP062966)
DE$lfcSE <- (DE$CI.R-DE$CI.L)/(2*1.96)
DE_analysis$SRP062966 <- DE
colnames(DE_analysis$SRP062966) <- str_c(colnames(DE_analysis$SRP062966),"_SRP062966")
write_csv(as.data.frame(DE_analysis$SRP062966), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP062966.csv")

DE <- do_limma(dge_3, rse_gene = rse_curated$SRP073191)
DE$lfcSE <- (DE$CI.R-DE$CI.L)/(2*1.96)
DE_analysis$SRP073191 <- DE
colnames(DE_analysis$SRP073191) <- str_c(colnames(DE_analysis$SRP073191),"_SRP073191")
write_csv(as.data.frame(DE_analysis$SRP073191), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP073191.csv")

DE <- do_limma(dge_4, rse_gene = rse_curated$SRP136102)
DE$lfcSE <- (DE$CI.R-DE$CI.L)/(2*1.96)
DE_analysis$SRP136102 <- DE
colnames(DE_analysis$SRP136102) <- str_c(colnames(DE_analysis$SRP136102),"_SRP136102")
write_csv(as.data.frame(DE_analysis$SRP136102), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP136102.csv")
```

DESeq2 for other studies

```{r}
DE <- do_DESeq(rse_curated$SRP111941)
DE$CI.L = DE$log2FoldChange - 1.96*DE$lfcSE
DE$CI.R = DE$log2FoldChange + 1.96*DE$lfcSE
DE_analysis$SRP111941 <- DE
colnames(DE_analysis$SRP111941) <- str_c(colnames(DE_analysis$SRP111941),"_SRP111941")
write_csv(as.data.frame(DE_analysis$SRP111941), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP111941.csv")

#DE_analysis$SRP062966 <- do_DESeq(rse_curated$SRP062966)
# write_csv(as.data.frame(DE_analysis$SRP062966), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP062966.csv")

DE <- do_DESeq(rse_curated$SRP311059)
DE$CI.L = DE$log2FoldChange - 1.96*DE$lfcSE
DE$CI.R = DE$log2FoldChange + 1.96*DE$lfcSE
DE_analysis$SRP311059 <- DE
colnames(DE_analysis$SRP311059) <- str_c(colnames(DE_analysis$SRP311059),"_SRP311059")
write_csv(as.data.frame(DE_analysis$SRP311059), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP311059.csv")

DE<- do_DESeq(rse_curated$SRP168421)
DE$CI.L = DE$log2FoldChange - 1.96*DE$lfcSE
DE$CI.R = DE$log2FoldChange + 1.96*DE$lfcSE
DE_analysis$SRP168421 <- DE
colnames(DE_analysis$SRP168421) <- str_c(colnames(DE_analysis$SRP168421),"_SRP168421")
write_csv(as.data.frame(DE_analysis$SRP168421), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP168421.csv")

DE <- do_DESeq(rse_curated$SRP322015)
DE$CI.L = DE$log2FoldChange - 1.96*DE$lfcSE
DE$CI.R = DE$log2FoldChange + 1.96*DE$lfcSE
DE_analysis$SRP322015 <- DE
colnames(DE_analysis$SRP322015) <- str_c(colnames(DE_analysis$SRP322015),"_SRP322015")
write_csv(as.data.frame(DE_analysis$SRP322015), file = "/mnt/Citosina/amedina/ssalazar/meta/out/DE2/Table_Genes_SRP322015.csv")
```

## Session info
```{r}
sessionInfo()
# R version 4.0.2 (2020-06-22)
# Platform: x86_64-pc-linux-gnu (64-bit)
# Running under: CentOS Linux 7 (Core)

# Matrix products: default
# BLAS:   /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRblas.so
# LAPACK: /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRlapack.so

# locale:
 # [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
 # [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
 # [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 # [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
 # [9] LC_ADDRESS=C               LC_TELEPHONE=C            
# [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

# attached base packages:
# [1] parallel  stats4    stats     graphics  grDevices utils     datasets 
# [8] methods   base     

# other attached packages:
 # [1] forcats_0.5.1               stringr_1.4.0              
 # [3] dplyr_1.0.3                 purrr_0.3.4                
 # [5] readr_1.4.0                 tidyr_1.1.2                
 # [7] tibble_3.1.3                ggplot2_3.3.5              
 # [9] tidyverse_1.3.0             edgeR_3.32.1               
# [11] limma_3.46.0                recount3_1.0.7             
# [13] SummarizedExperiment_1.20.0 Biobase_2.50.0             
# [15] GenomicRanges_1.42.0        GenomeInfoDb_1.26.7        
# [17] IRanges_2.24.1              S4Vectors_0.28.1           
# [19] BiocGenerics_0.36.1         MatrixGenerics_1.2.1       
# [21] matrixStats_0.60.0         

# loaded via a namespace (and not attached):
 # [1] httr_1.4.2               jsonlite_1.7.2           bit64_4.0.5             
 # [4] R.utils_2.10.1           modelr_0.1.8             assertthat_0.2.1        
 # [7] BiocFileCache_1.14.0     blob_1.2.1               cellranger_1.1.0        
# [10] GenomeInfoDbData_1.2.4   Rsamtools_2.6.0          sessioninfo_1.1.1       
# [13] pillar_1.6.2             RSQLite_2.2.7            backports_1.2.1         
# [16] lattice_0.20-41          glue_1.4.2               XVector_0.30.0          
# [19] rvest_0.3.6              colorspace_2.0-2         Matrix_1.3-4            
# [22] R.oo_1.24.0              XML_3.99-0.6             pkgconfig_2.0.3         
# [25] broom_0.7.9              haven_2.3.1              zlibbioc_1.36.0         
# [28] scales_1.1.1             BiocParallel_1.24.1      generics_0.1.0          
# [31] ellipsis_0.3.2           cachem_1.0.5             withr_2.4.2             
# [34] cli_3.0.1                readxl_1.3.1             magrittr_2.0.1          
# [37] crayon_1.4.1             memoise_2.0.0            R.methodsS3_1.8.1       
# [40] fs_1.5.0                 fansi_0.5.0              xml2_1.3.2              
# [43] tools_4.0.2              data.table_1.14.0        hms_1.1.0               
# [46] lifecycle_1.0.0          reprex_1.0.0             munsell_0.5.0           
# [49] locfit_1.5-9.4           DelayedArray_0.16.3      Biostrings_2.58.0       
# [52] compiler_4.0.2           rlang_0.4.11             grid_4.0.2              
# [55] RCurl_1.98-1.3           rstudioapi_0.13          rappdirs_0.3.3          
# [58] bitops_1.0-7             gtable_0.3.0             DBI_1.1.1               
# [61] curl_4.3.2               R6_2.5.0                 lubridate_1.7.9.2       
# [64] GenomicAlignments_1.26.0 rtracklayer_1.50.0       fastmap_1.1.0           
# [67] bit_4.0.4                utf8_1.2.2               stringi_1.6.2           
# [70] Rcpp_1.0.7               vctrs_0.3.8              dbplyr_2.0.0            
# [73] tidyselect_1.1.0  
```




