---
title: "meta-analysis"
author: "Sofia Salazar"
date: "5/17/2022"
output: html_document
---
## Setup

This script is used to perform a Meta-analysis of the studies in the `SRA-IDs.txt` file.

```{bash}
qlogin
cd /mnt/Citosina/amedina/ssalazar/meta/out
module load r/4.0.2
module load pandoc/2.9.2
R
```

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(ComplexHeatmap)
library(metafor)
library(pheatmap)
library(RColorBrewer)
library(meta)
library("MetaVolcanoR")
library(ggrepel)
library("biomaRt")
```

## Read data

```{r}
full_tables <- read_csv("/mnt/Citosina/amedina/ssalazar/meta/out/fullTables.csv")
colnames(full_tables)
rse_curated <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/curated_rse.xz")
SRA_info <- read_csv("/mnt/Citosina/amedina/ssalazar/meta/out/summary.csv")
```


Filtering genes by selecting only protein coding genes in the human genome using biomaRt.

```{r}
mart <- useMart(biomart="ensembl", dataset="hsapiens_gene_ensembl")
all_genes <- getBM(attributes = c( "hgnc_symbol","transcript_biotype"), 
                          filters = c("hgnc_symbol"), 
                          values = full_tables$gene_name,
                          #values = list(biotype=c("protein_coding")), 
                          mart = mart)
genes_emb <- full_tables$gene_name %in% all_genes$hgnc_symbol
```

## p-Values

We first check pvalues histograms. [Ref](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6164648/)

```{r}
outdir = "/mnt/Citosina/amedina/ssalazar/meta/out/"
FC <- str_subset(colnames(full_tables),"logFC")
# Genes with a adjusted p value
adjp <- full_tables %>%
  filter(gene_name != "NA") %>%
  gather("SRA"  ,"adjP",-gene_name) %>%
  filter(SRA %in% str_subset(colnames(full_tables),"adj.P.Val|padj")) %>%
  mutate(SRA=gsub(".*_","",SRA))
  
adjp
#pdf("Pvalues.pdf",width = 8,height = 6)
png(file = paste0(outdir,"adj-pvalue-histogram.png"), width = 800, height = 800)
adjp %>%    
  group_by(SRA) %>%
  ggplot( aes(x=adjP, fill=SRA)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0.05, linetype="dashed", color = "red") +
  theme_bw(16) +
  xlab("adjusted p-value") +
  ylab("Frequency") +
  facet_wrap(~SRA, scales = "free_y") 
dev.off()
```

Then, we look for how many genes are up regulated with a p-value less than `r  adj_pvalue`  in the 

```{r}
# log fold change
logFC <- full_tables %>%
  filter(gene_name != "NA") %>%
  gather("SRA"  ,"logFC",-gene_name) %>%
  filter(SRA %in% str_subset(colnames(full_tables),"logFC")) %>% 
  mutate(SRA=gsub(".*_","",SRA))
logFC
png(file = paste0(outdir,"logFC-histogram.png"), width = 800, height = 800)
logFC %>%    
  group_by(SRA) %>%
  ggplot( aes(x=logFC, fill=SRA)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0, linetype="dashed", color = "red") +
  theme_bw(16) +
  xlab("logFC") +
  ylab("Frequency") +
  facet_wrap(~SRA, scales = "free_x") 
dev.off()
# recovering logFC values from genes with p value less than 0.01 
adj_pvalue <- 0.05
ap <- left_join(adjp,logFC,by = c("gene_name" = "gene_name", "SRA" = "SRA")) %>%
  filter(adjP < adj_pvalue) %>% 
  dplyr::select(-adjP) %>% 
  spread(SRA,logFC) %>% 
  replace(is.na(.), 0)
ap
# Experiments with no DE genes 
#unique(logFC$SRA)[which(!unique(logFC$SRA)  %in% colnames(ap))]
#ap$GSE42057 <- 0
# selecting up regulated genes
up <- ap %>%
  mutate_at(vars(-gene_name),list(~if_else( .>0, 1, 0)))
  
down <- ap %>%
  mutate_at(vars(-gene_name),list(~if_else( .<0, 1, 0)))

### UPREGULATED GENES

m <- make_comb_mat(up,mode = "distinct")
#m <- m[comb_degree(m) >= 1]
m <- m[comb_degree(m) >= 2 | comb_degree(m) == 1]
#pdf("upsetUP.pdf",width = 14,height = 6)

png(file = paste0(outdir,"UpSet-plot-upregulated.png"), width = 800, height = 800)
ht <- draw(UpSet(m,
      pt_size = unit(5, "mm"), 
      lwd = 3,
      comb_col = c("red", "#5c5c5c","#5c5c5c","#5c5c5c", "black","blue")[comb_degree(m)])
      #order.by = "freq"
    )
od = column_order(ht)
cs = comb_size(m)
decorate_annotation("intersection_size", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 12))
})
dev.off()

comb_size(m)
sample(up$gene_name[extract_comb(m,"11")],)
q
### DOWNREGULATED GENES

m <- make_comb_mat(down,mode = "distinct")
#m <- m[comb_degree(m) >= 1]
m <- m[comb_degree(m) >= 2 | comb_degree(m) == 1] 
#pdf("upsetDOWN.pdf",width = 14,height = 6)
png(file = paste0(outdir,"UpSet-plot-downregulated.png"), width = 800, height = 800)
ht <- draw(UpSet(m,
      pt_size = unit(4, "mm"), 
      lwd = 3,
      comb_col = c("blue", "#5c5c5c","#5c5c5c","#5c5c5c", "black","red")[comb_degree(m)])
      #order.by = "freq"
    )
od = column_order(ht)
cs = comb_size(m)
decorate_annotation("intersection_size", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 12))
})
dev.off()
comb_size(m)
sample(up$gene_name[extract_comb(m,"11")],)
```

## Meta-analysis for a single gene

We first calculate the Standar error using confidence intervals

```{r}
CI.R <- str_subset(colnames(full_tables),"CI.R")
CI.L <- str_subset(colnames(full_tables),"CI.L")
for(i in 1:length(CI.R)) {
  full_tables[str_c("SE_",gsub(".*_","",CI.R))] <- (full_tables[,CI.R]-full_tables[,CI.L]) /(2*1.96)
}
```

```{r}
FC <- str_subset(colnames(full_tables),"logFC")
SE <- str_subset(colnames(full_tables),"SE_")
SRA <- gsub(".*_","",FC)
## Function to perform a meta-analysis per gene
meta.plot <- function(gs){
  gene <-  full_tables[full_tables$gene_name == gs,] # IFI44L
  #gene <-  na.omit(full_tables[full_tables$gene_name == gs,])
  gene <- gene[!is.na(gene$gene_name),]
  dat <- data.frame(
    yi = as.numeric(gene[,FC]), # effect sizes observed (Fold change)
    sei = as.numeric(gene[,SE]),
    n1i = as.numeric(SRA_info$CONTROL),
    n2i = as.numeric(SRA_info$SLE),
    study = SRA_info$SRA
  )

  dat <- escalc(measure="ZCOR", yi=yi,sei=sei,n1i=n1i,n2i=n2i, data = dat) 
### meta analysis in the simplest way
#png(file= paste0(outdir,"forest-test.png"), width = 800, height = 800)
  #res <- metafor::rma(yi, sei=sei,n1i=n1i,n2i=n2i, data=dat)
  #funnel(res)
  #forest(res,slab = SRA, main=unique(gene$gene_name))
#dev.off()
#####
  dat$study <- SRA_info$SRA
  m1 <- metagen(TE = yi, 
                  seTE = sei, 
                  studlab = study, 
                  n.e = n2i, 
                  n.c = n1i, 
                  data = dat,
                  method.tau = "REML", 
                  prediction = TRUE,
                  sm= "MD",
                  fixed = FALSE)
  forest(m1,
       col.diamond = "blue",
       leftcols = c("study"),
       #rightcols = c("ci", "w.random"),
       leftlabs =  c("Study","logFC","se",NA,NA),
       colgap.left = unit(5,"mm"),
       label.e = "SLE",
       colgap.studlab = unit(-10,"mm"),
       ff.study.label = "bold",
       just = "center",
       xlab = gs,
       prediction = TRUE)
}

```

## Meta-analysis for all genes

Saving the results in a new dataframe.

```{r}
#gs <- unique(na.omit(full_tables)$GENE.SYMBOL)
gs <- full_tables$gene_name
meta <- data.frame(genes = 0,
                   TE.random = 0,
                   pval.random = 0,
                   tau2 = 0,
                   I2 = 0,
                   pval.Q = 0,
                   num_exp=0)
# meta <- data.frame(genes = 0,
#                    QEp=0,
#                    estimate=0,
#                    pval=0,
#                    se=0,
#                    H2=0,
#                    tau2=0,
#                    I2=0,
#                    num_exp=0)
gm <- full_tables$gene_name
full_tables[full_tables == 0] <- NA
length(gm)
for (i in gm) { 
#for( i in gm[18705:length(gm)]){ #problemas con 
  gene <- full_tables[which(full_tables$gene_name == i),]
  dat <- data.frame(
    yi = as.numeric(gene[,FC]),
    sei = as.numeric(gene[,SE]),
    n1i = as.numeric(SRA_info$CONTROL),
    n2i = as.numeric(SRA_info$SLE)
  )
  dat$study <- rownames(SRA_info)
  #dat$country <- gse_table$COUNTRY
  
  # meta analysis with random models "REML"
  # https://www.rdocumentation.org/packages/meta/versions/4.9-6/topics/metagen
  m1 <- metagen(TE = yi, 
                  seTE = sei, 
                  studlab = study, 
                  n.e = n2i, 
                  n.c = n1i, 
                  data = dat,
                  method.tau = "REML", 
                  prediction = TRUE,
                  sm= "MD",
                  control=list(maxiter=1000), # increasing number of iteration
                  comb.fixed = FALSE)
  # res <- suppressWarnings(metafor::rma(yi, sei=sei,control=list(stepadj=.5,maxiter=1000)))
  num_exp <- sum(!is.na(dat$yi))
  meta <- rbind(meta,
                c(i, m1$TE.random,m1$pval.random,m1$tau2,m1$I2,m1$pval.Q,num_exp))
  #meta <- rbind(meta,c(i,res$QEp,res$beta,res$pval,res$se,res$H2,res$tau2,res$I2,num_exp))
}
# Numeric values
meta <- meta[-1,]
rownames(meta) <- meta$genes
meta <- meta[,-1] %>% mutate_all(as.numeric)
# FDR correction
meta$qval.random <- p.adjust(meta$pval.random,method="fdr")
meta$qval.Q <- p.adjust(meta$pval.Q, method = "fdr")
```

### Output of full genes meta-analysis
```{r}
write.csv(meta, file ="/mnt/Citosina/amedina/ssalazar/meta/out/meta-analysis-6.csv")
```

```{r}
meta <- read_csv("/mnt/Citosina/amedina/ssalazar/meta/out/meta-analysis-6.csv")
### Distribution of statistics 
png(file= paste0(outdir,"distribution-stats-full.png"), width = 800, height = 800)
  bqval <- ggplot(meta) +
    aes(x = qval.random) +
    geom_histogram(bins = 55L, fill = "#c45a75") +
    labs(x = "q-value from REML") +
    theme_classic()
  bq <- ggplot(meta) +
    aes(x = qval.Q) +
    geom_histogram(bins = 55L, fill = "#c45a75") +
    labs(x = "q-value for Q test") +
    theme_classic()
  bte <- ggplot(meta) +
    aes(x = TE.random) +
    geom_histogram(bins = 55L, fill = "#c45a75") +
    labs(x = "Estimate values") +
    theme_classic()
  btau <- ggplot(meta) +
    aes(x = tau2) +
    geom_histogram(bins = 55L, fill = "#c45a75") +
    labs(x = "Tau2") +
    theme_classic()
  bi2 <- ggplot(meta) +
    aes(x = I2) +
    geom_histogram(bins = 55L, fill = "#c45a75") +
    labs(x = "I2") +
    theme_classic()
  ggpubr::ggarrange(bqval,bq,bte,btau,bi2,
                  nrow =2,
                  ncol =3)
dev.off()
### Selecting genes
qval_cutoff <- 0.05
genesqval <- meta %>% 
  filter(meta$qval.random <= qval_cutoff & I2 < 0.40 & num_exp > 1) %>% 
  arrange(TE.random)
genesqval
kable(c(head(genesqval),tail(genesqval)), caption = "Meta-analysis")

### logFC values from selected genes
topgenes <- full_tables %>% 
  filter(full_tables$gene_name %in% genesqval$X1) %>% 
  dplyr::select(gene_name,contains("logFC"))
## as dataframe
topgenes <- column_to_rownames(topgenes,"gene_name")
# scaling data
topgenes <- scale(topgenes)
#"BrBG"
color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))
```

```{r fig.height=15}
dim(topgenes)
png(file= paste0(outdir,"heatmap-topgenes-qval.png"), width = 800, height = 800)
  pheatmap(topgenes,
          main = str_c("Genes with q-value less than ",qval_cutoff, " (ngenes = ",dim(topgenes)[1], ")"),
          color = color,
          breaks = breaks,
          fontsize_row = 7,
          fontsize_col = 10,
          cluster_rows = T,
          clustering_distance_rows = "euclidean",
          clustering_method = "ward.D",
          cluster_cols = T,
          cutree_rows = 2,
          show_colnames = T,
          show_rownames = F,
          border_color = NA)
dev.off()
```



```{r}
mo <-  genesqval %>% 
  arrange(TE.random)
mo <- rbind(head(mo,15), tail(mo,15))
### logFC values from selected genes
topgenes <- full_tables %>% 
  filter(full_tables$gene_name %in% mo$X1) %>% 
  dplyr::select(gene_name,contains("logFC")) %>% 
  column_to_rownames(.,"gene_name") %>%
  na.omit(.)
topgenes <- scale(topgenes)
color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))

png(file= paste0(outdir,"heatmap-logFC-and-qval.png"))
  pheatmap(topgenes,
         main = str_c("The highest logFC and qval < " ,qval_cutoff , " (ngenes = ",dim(topgenes)[1], ")"),
         color = color,
         breaks = breaks,
         fontsize_row = 9,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 2,
         show_colnames = T,
         border_color = NA)
dev.off()
```

### MetaVolcano

Using [Metavolcano](https://www.bioconductor.org/packages/release/bioc/vignettes/MetaVolcanoR/inst/doc/MetaVolcano.html) funcions to create a volcano plot with DE genes 

We first need to create a list with tables. Each table contains information about gene `Symbol`, `Log2FC`, `pvalue`, `CI.L`, `CI.R` per experiment. 

```{r}
# colnames 
df <- tibble(x=colnames(full_tables)) %>% 
  separate(x, c("type", "SRA"), "_") %>% 
  add_column(pos = 1:nrow(.))
df$SRA[1] <- "Gene"
# list
split_df <- function(i){
  df <- data.frame(gene_name = full_tables$gene_name[genes_emb],
                   full_tables[genes_emb,which(df$SRA == i)]
                   )
  colnames(df) <- c( "Symbol", "Log2FC",  "pvalue", "CI.L", "CI.R")
  #colnames(df) <- gsub("_GSE.*", "", colnames(df) )
  as.data.frame(df)
  #as.data.frame(df)[-c(8359,18704),]
  }
diffexplist <- lapply(sort(unique(df$SRA)[-1]), split_df)
names(diffexplist) <- sort(unique(df$SRA)[-1])
```



```{r}
plot_rem <- function(meta_diffexp, jobname, outputfolder, genecol, metathr) {
    meta_diffexp$randomPadj <- p.adjust(meta_diffexp$randomP, "BH")
    irandomp <- 0.05
    meta_diffexp %>%
        dplyr::mutate(signcon2 = ifelse(`randomP` <= irandomp, signcon, NA)) %>%
	dplyr::mutate(Ci.ub = ifelse(`randomP` <= irandomp, randomCi.ub, NA)) %>%
	dplyr::mutate(Ci.lb = ifelse(`randomP` <= irandomp, randomCi.lb, NA)) %>%
	dplyr::filter(`randomP` <  quantile(meta_diffexp[['randomP']], 0.6)) -> meta_res 
    ggplot(dplyr::arrange(meta_res, abs(randomSummary)),
        aes(x = randomSummary, y = -log10(randomPadj), color = signcon2, 
	    text = !!rlang::sym(genecol))) +
        geom_point(size = 0.6) +
	scale_color_gradient2(midpoint=0, low="#195696", mid="white", 
			      high="red", na.value = "grey80") +
	labs(x = "Summary Fold-change",
	     y = "-log10(Summary padj-value)",
	     color = "Sign consistency") +
        geom_errorbarh(aes(xmax = Ci.ub, xmin = Ci.lb, 
			   color = signcon2)) +
        theme_classic(14) +
	theme(panel.border= element_blank()) +
	theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
	theme(axis.line.x = element_line(color = "black", size = 0.6, 
					 lineend = "square"),
	      axis.line.y = element_line(color = "black", size = 0.6, 
					 lineend = "square"))
}
```

Now, we implement the `MetaVolcano` funtions. 

```{r}
meta_degs_rem <- rem_mv(diffexp=diffexplist,
            pcriteria="pvalue",
            foldchangecol='Log2FC', 
            genenamecol='Symbol',
            geneidcol=NULL,
            collaps=FALSE,
            llcol='CI.L',
            rlcol='CI.R',
            vcol=NULL, 
            cvar=TRUE,
            metathr=0.01,
            jobname="MetaVolcano",
            outputfolder=".",
            draw='HTML',
            ncores=1)
meta_res <- meta_degs_rem@metaresult
top <- head(meta_degs_rem@metaresult[order(meta_degs_rem@metaresult$randomP),], 30)
top <- head(meta_degs_rem@metaresult, 3)
meta_degs_rem@MetaVolcano <- plot_rem(meta_res, 
               outputfolder="/mnt/Citosina/amedina/ssalazar/meta/out",
               genecol = 'Symbol', 
               metathr=0.01) +
  #geom_label_repel(aes(label = ifelse(Symbol %in% top$Symbol, Symbol, "")),
  geom_label_repel(aes(label = ifelse(Symbol %in% rownames(topgenes), Symbol, "")),
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  size = 4)
pdf(file= paste0(outdir,"metavolcano.pdf"), width = 20, height = 20)
  meta_degs_rem@MetaVolcano
dev.off()
meta_df <- meta_degs_rem@metaresult %>% 
  filter(randomP < 0.05 )
```

Forest plot
```{r}
png(file= paste0(outdir,"forestplot-IFIT1.png"), width = 800, height = 800)
  meta.plot("IFIT1")
dev.off()
png(file= paste0(outdir,"forestplot-IFI6.png"), width = 800, height = 800)
  meta.plot("IFI6")
dev.off()
png(file= paste0(outdir,"forestplot-IFIT3.png"), width = 800, height = 800)
  meta.plot("IFIT3")
dev.off()
```

## Vote counting approach

```{r}
# vote counting method
meta_degs_vote <- votecount_mv(diffexp=diffexplist,
                   pcriteria='pvalue',
                   foldchangecol='Log2FC',
                   genenamecol='Symbol',
                   geneidcol=NULL,
                   pvalue=0.05,
                   foldchange=0, 
                   metathr=0.01,
                   collaps=FALSE,
                   jobname="MetaVolcano", 
                   outputfolder=".",
                   draw='HTML')
meta_degs_vote <- votecount_mv(diffexp=diffexplist,
            pcriteria="pvalue",
            foldchangecol='Log2FC', 
            genenamecol='Symbol',
            geneidcol=NULL,
            collaps=FALSE,
            #llcol='CI.L',
            #rlcol='CI.R',
            #vcol=NULL, 
            #cvar=TRUE,
            metathr=0.05,
            jobname="MetaVolcano",
            outputfolder=".",
            draw='HTML')
            #ncores=1)
head(meta_degs_vote@metaresult, 3)
meta_degs_vote@degfreq
meta_degs_vote@MetaVolcano
```