---
title: "normalization"
author: "Sofia Salazar"
date: "5/16/2022"
output: html_document
---

## Setup

In this script we use `edgeR` for normalization by creating a `DGEList` object for each `rse` objects obtained from the previous `curation-2.Rmd` script.

```{bash}
qlogin
cd /mnt/Citosina/amedina/ssalazar/meta/out
module load r/4.0.2
R
```

## Libraries

```{r}
library("recount3")
library("edgeR")
library(tidyverse)
library("sva")
library(tidyverse)
library("SummarizedExperiment")
library(limma)
```



## Normalization

Getting the curated rse objects obtained from previous `curation.Rmd`.

```{r}
SRA_projects <- scan(file = "/mnt/Citosina/amedina/ssalazar/meta/data/SRA-IDs.txt", character(), sep=",")
info <- read_csv(file = "/mnt/Citosina/amedina/ssalazar/meta/out/summary.csv")
curated <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/corrected_rse.xz")
rse <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/curated_rse.xz")
normalization <- readRDS("/mnt/Citosina/amedina/ssalazar/meta/out/corrected_rse.xz")
```

## PCA plots pre-normalization

```{r}
for (i in 1:length(curated)){
  png(file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/figures/PCAs/PCpre", names(curated)[i], ".png"), width = 800, height = 800)

  plotMDS(curated[[i]], labels = curated[[i]]$DISEASE, col=ifelse(curated[[i]]$DISEASE=="CONTROL","blue","red"), top=200, gene.selection="common", prior.count = 5)
  dev.off()
}
```


Function for creating a `dge` object for each `rse` object.

```{r}
get_dge <- function(rse){
  dge <- DGEList(counts = assay(rse, "counts"), genes = rowData(rse))
  # Filtering out rows with zero or very low counts
  design <- model.matrix(~ DISEASE, data = colData(rse))
  keep <- filterByExpr(dge, design)
  dge <- dge[keep,,keep.lib.sizes=FALSE]
  # normalizing
  dge <- calcNormFactors(dge)
  return(dge)
}
```


Performing normalization 

## Batch correction 

*for SRP168421*

```{r}
rse_object <- curated$SRP168421
batch<- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,1,1)
corrected <- ComBat_seq(counts = assays(rse_object)$counts, batch = batch, group = rse_object$DISEASE)
assays(curated$SRP168421)$counts <- corrected
# pca FOR batch corrected
png(file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/figures/PCAs/PCpre-correctedSRP16.png"), width = 800, height = 800)
  plotMDS(curated$SRP168421, labels = curated$SRP168421$DISEASE, col=ifelse(curated$SRP168421$DISEASE=="CONTROL","blue","red"), top=200, gene.selection="common", prior.count = 5)
  dev.off()
assays(rse$SRP168421)$counts <- corrected

# normalizing
dge_result <- get_dge(rse_object)
curated$SRP168421 <- dge_result
design <- model.matrix(as.formula(~ rse_object$DISEASE), data = colData(rse_object))
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP168421")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP168421.csv"))
```

*for SRP073191*

```{r}
rse_object <- curated$SRP073191
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP073191.tsv", header = T, sep= ',')
pheno <- sra_info[, c(1, 13)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(disease_state), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 3
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP073191 <- design

# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP073191")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP073191.csv"))
```

*for SRP062966*

```{r}
rse_object <- curated$SRP062966
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP062966.tsv", header = T, sep= '\t')
pheno <- sra_info[, c(1, 14)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(disease_status), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 13
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP062966 <- design
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP062966")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP062966.csv"))
```

*for SRP136102*

```{r}
rse_object <- curated$SRP136102
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP136102.tsv", header = T, sep= ',')
pheno <- sra_info[, c(1, 24)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(phenotype), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 16
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP136102 <- design
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP136102")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP136102.csv"))
```

*for SRP111941*

```{r}
rse_object <- curated$SRP111941
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP111941.tsv", header = T, sep= '\t')
pheno <- sra_info[, c(1, 2)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(subject_status), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 2
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP111941 <- design
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP111941")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP111941.csv"))
```

*for SRP296987*

```{r}
rse_object <- curated$SRP296987
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP296987.tsv", header = T, sep= '\t')
pheno <- sra_info[, c(1, 2)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(source_name), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 2
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP296987 <- design
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP296987")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP296987.csv"))
```

*for SRP311059*

```{r}
rse_object <- curated$SRP311059
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP311059.tsv", header = T, sep= '\t')
pheno <- sra_info[, c(1, 2)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(Diagnosis), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 4
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP311059 <- design
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP311059")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP311059.csv"))
```

*for SRP322015* 

```{r}
rse_object <- curated$SRP322015
sra_info <- read.delim(file = "/mnt/Citosina/amedina/ssalazar/meta/data/info-SRP322015.tsv", header = T, sep= '\t')
pheno <- sra_info[, c(1, 2)] # phenotype info
# normalizing
dge_result <- get_dge(rse_object)
mod <- model.matrix(~as.factor(source_name), data=pheno)
mod0 <- model.matrix(~1,data=pheno) # null model
svobj <- svaseq(cpm(dge_result), mod, mod0) # significant surrogate variables 2
design <- cbind(mod, svobj$sv) # design matrix
curated$SRP322015 <- design
# fitting limma 
vGene <- voom(dge_result, design, plot = F)
eb_results <- eBayes(lmFit(vGene))
de_results <- topTable(eb_results, coef = ncol(design), number = nrow(rse_object), sort.by = "none", confint = T)
res <- as_tibble(de_results,rownames="rownames")
colnames(res) <- str_c(colnames(res),"_SRP322015")
write_csv(as.data.frame(res), file = str_c("/mnt/Citosina/amedina/ssalazar/meta/out/batch/Table_Genes_SRP322015.csv"))
```

## output for batch correction
```{r}
saveRDS(curated, file = "/mnt/Citosina/amedina/ssalazar/meta/out/batch_all_dge.xz", compress = "xz")
```


## limma method for all studies

```{r}
for(i in 1:length(SRA_projects)){
  rse_object <- curated[i]
  # apply the function to the rse object inside each list name
  dge_result <- lapply(rse_object, get_dge)
  # overwriting each list element for the dge object
  curated[i] <- dge_result
}
saveRDS(curated, file = "/mnt/Citosina/amedina/ssalazar/meta/out/all_dge.xz", compress = "xz")
```


## Different methods for each study

For study SRP296987

```{r}
#for(i in 1:length(SRA_projects)){
  rse_object <- curated$SRP296987
  # apply the function to the rse object inside each list name
  # dge_result <- lapply(rse_object, get_dge)
 dge_result <- get_dge(rse_object)
 saveRDS(dge_result, file = "/mnt/Citosina/amedina/ssalazar/meta/out/dge_mix/SRP296987_dge.xz", compress = "xz")
```

For SRP062966
```{r}
rse_object <- curated$SRP062966
dge_result <- get_dge(rse_object)
saveRDS(dge_result, file = "/mnt/Citosina/amedina/ssalazar/meta/out/dge_mix/SRP062966_dge.xz", compress = "xz")
```

For SRP073191

```{r}
rse_object <- curated$SRP073191
dge_result <- get_dge(rse_object)
saveRDS(dge_result, file = "/mnt/Citosina/amedina/ssalazar/meta/out/dge_mix/SRP073191_dge.xz", compress = "xz")
```

For SRP136102

```{r}
rse_object <- curated$SRP136102
dge_result <- get_dge(rse_object)
saveRDS(dge_result, file = "/mnt/Citosina/amedina/ssalazar/meta/out/dge_mix/SRP136102_dge.xz", compress = "xz")
```

## Batch correction for SRP168421

```{r}
rse_object <- curated$SRP168421
batch<- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,1,1)
corrected <- ComBat_seq(counts = assays(rse_object)$counts, batch = batch, group = rse_object$DISEASE)
assays(rse_object)$counts <- corrected
curated$SRP168421 <- rse_object
```

## Output

The output is a list of the dge objects obtained for each study.

```{r}
# saveRDS(normalization, file = "/mnt/Citosina/amedina/ssalazar/meta/out/normalized_dge.xz", compress = "xz")
saveRDS(rse, file = "/mnt/Citosina/amedina/ssalazar/meta/out/corrected_rse.xz")
```

## Session info

```{r}
sessionInfo()
# R version 4.0.2 (2020-06-22)
# Platform: x86_64-pc-linux-gnu (64-bit)
# Running under: CentOS Linux 7 (Core)

# Matrix products: default
# BLAS:   /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRblas.so
# LAPACK: /cm/shared/apps/r/4.0.2-studio/lib64/R/lib/libRlapack.so

# locale:
 # [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
 # [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
 # [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
 # [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
 # [9] LC_ADDRESS=C               LC_TELEPHONE=C            
# [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       

# attached base packages:
# [1] parallel  stats4    stats     graphics  grDevices utils     datasets 
# [8] methods   base     

# other attached packages:
 # [1] sva_3.38.0                  BiocParallel_1.24.1        
 # [3] genefilter_1.72.1           mgcv_1.8-33                
 # [5] nlme_3.1-151                forcats_0.5.1              
 # [7] stringr_1.4.0               dplyr_1.0.3                
 # [9] purrr_0.3.4                 readr_1.4.0                
# [11] tidyr_1.1.2                 tibble_3.1.3               
# [13] ggplot2_3.3.5               tidyverse_1.3.0            
# [15] edgeR_3.32.1                limma_3.46.0               
# [17] recount3_1.0.7              SummarizedExperiment_1.20.0
# [19] Biobase_2.50.0              GenomicRanges_1.42.0       
# [21] GenomeInfoDb_1.26.7         IRanges_2.24.1             
# [23] S4Vectors_0.28.1            BiocGenerics_0.36.1        
# [25] MatrixGenerics_1.2.1        matrixStats_0.60.0         

# loaded via a namespace (and not attached):
 # [1] bitops_1.0-7             fs_1.5.0                 lubridate_1.7.9.2       
 # [4] bit64_4.0.5              httr_1.4.2               tools_4.0.2             
 # [7] backports_1.2.1          utf8_1.2.2               R6_2.5.0                
# [10] DBI_1.1.1                colorspace_2.0-2         withr_2.4.2             
# [13] tidyselect_1.1.0         bit_4.0.4                curl_4.3.2              
# [16] compiler_4.0.2           cli_3.0.1                rvest_0.3.6             
# [19] xml2_1.3.2               DelayedArray_0.16.3      rtracklayer_1.50.0      
# [22] scales_1.1.1             rappdirs_0.3.3           Rsamtools_2.6.0         
# [25] R.utils_2.10.1           XVector_0.30.0           pkgconfig_2.0.3         
# [28] sessioninfo_1.1.1        dbplyr_2.0.0             fastmap_1.1.0           
# [31] rlang_0.4.11             readxl_1.3.1             rstudioapi_0.13         
# [34] RSQLite_2.2.7            generics_0.1.0           jsonlite_1.7.2          
# [37] R.oo_1.24.0              RCurl_1.98-1.3           magrittr_2.0.1          
# [40] GenomeInfoDbData_1.2.4   Matrix_1.3-4             Rcpp_1.0.7              
# [43] munsell_0.5.0            fansi_0.5.0              lifecycle_1.0.0         
# [46] R.methodsS3_1.8.1        stringi_1.6.2            zlibbioc_1.36.0         
# [49] BiocFileCache_1.14.0     grid_4.0.2               blob_1.2.1              
# [52] crayon_1.4.1             lattice_0.20-41          splines_4.0.2           
# [55] Biostrings_2.58.0        haven_2.3.1              annotate_1.68.0         
# [58] hms_1.1.0                locfit_1.5-9.4           pillar_1.6.2            
# [61] reprex_1.0.0             XML_3.99-0.6             glue_1.4.2              
# [64] data.table_1.14.0        modelr_0.1.8             vctrs_0.3.8             
# [67] cellranger_1.1.0         gtable_0.3.0             assertthat_0.2.1        
# [70] cachem_1.0.5             xtable_1.8-4             broom_0.7.9             
# [73] survival_3.2-7           AnnotationDbi_1.52.0     GenomicAlignments_1.26.0
# [76] memoise_2.0.0            ellipsis_0.3.2
```



